Hemispheric Lateralization and Outcomes After ICH: A Pooled Analysis of Clinical Trials

### 2022-11-16 study protocol.qmd
Update
More updates

library(tidyverse)
library(tableone)
library(data.table)
library(haven) # Importing SAS Files
library(janitor)
library(lubridate)

@@

**Importing and wrangling ATACH2 data

@@

atach.whole <- atach.whole %>%
  mutate(trach = recode(F07Q11,
                                   "0" = "No",
                                   "1" = "Yes",
                                   .default = "NA"))

@@

VarsTableOne <- c("study",
                  "age", # Age
                       "sex", # Sex
                       "ethnicity", # Ethnicity
                       "race", # Race
                       "enrolling.country", # Country of Enrolling Site
                       "stroke", 
                       "cad", 
                       "chf", 
                       "afib", 
                       "pvd",
                       "hld",
                       "dm2",
                       "tobacco",
                       #"symptom.onset", # Time of symptom onset
                       "sbp.initial", # Initial SBP
                       "dbp.initial", # Initial DBP
                       "gcs.baseline", # Baseline GCS
                       #"nihss.baseline", # NIHSS
                       "ich.location.baseline", 
                       "ivh",
                      # "hydrocephalus",
                       "ich.volume.cc",
                       "ivh.volume.cc",
                       "phe.volume.cc",
                       "sp.shift.baseline",
                       "mech.ventilation",
                       "intubation.days", 
                       "extubation.date", # Date of extubation
                       "trach",
                       "trach.date", # Date of Tracheostomy
                       "evd",
                       "evd.days", # Total EVD days
                       "nsgy.evacuation",
                       "nsgy.evacuation.date",  # Date of Evacuation
                       "mrs.30",
                       "mrs.90",
                      "mrs.180",
                      "mrs.365",
                      "euro.mobility.90",
                      "euro.mobility.180",
                      "euro.mobility.365",
                       "euro.selfcare.90",
                      "euro.selfcare.180",
                      "euro.selfcare.365",
                       "euro.usual.90",
                      "euro.usual.180",
                      "euro.usual.365",
                       "euro.pain.90",
                      "euro.pain.180",
                      "euro.pain.365",
                       "euro.anxiety.90",
                      "euro.anxiety.180",
                      "euro.anxiety.365",
                       "euro.vas.90", # EuroQOL VAS
                       "euro.vas.180",
                       "euro.vas.365",
                       "EQ_INDEX.90", # EQ-5D Utility Index,,
                      "current.location.90",
                      "current.location.180",
                      "current.location.365",
                       "ich.laterality"
                  )

# Maybe import lab data, but this may be superfluous

@@

Importing and wrangling ERICH data

@@

erich.whole <- erich.whole %>%
  mutate(evd = case_when(
    IT8 == "1" ~ "Yes",
    IT8 == "2" ~ "No",
  ))

erich.whole <- erich.whole %>%
  mutate(nsgy.evacuation = case_when(
    IT5D == "1" ~ "Yes",
    IT5H == "1" ~ "Yes",
    TRUE ~ "No"
      ))

@@

# Variables missing: nihss.baseline, hydrocephalus, phe.volume.cc, sp.shift.baseline, intubation.days, tracheostomy

@@

Combining Datasets

>>>>

library(tidyverse)
library(tableone)
library(data.table)
library(haven) # Importing SAS Files
library(janitor)
library(lubridate)
library(readxl)

@@

**Importing and wrangling ATACH2 data**

@@

atach.whole <- atach.whole %>%
  mutate(tracheostomy = recode(F07Q11,
                                   "0" = "No",
                                   "1" = "Yes",
                                   .default = "NA"))

@@

VarsTableOne <- c("study",
                  "age", # Age
                       "sex", # Sex
                       "ethnicity", # Ethnicity
                       "race", # Race
                       "enrolling.country", # Country of Enrolling Site
                       "stroke", 
                       "cad", 
                       "chf", 
                       "afib", 
                       "pvd",
                       "hld",
                       "dm2",
                       "tobacco",
                       #"symptom.onset", # Time of symptom onset
                       "sbp.initial", # Initial SBP
                       "dbp.initial", # Initial DBP
                       "gcs.baseline", # Baseline GCS
                       #"nihss.baseline", # NIHSS
                       "ich.location.baseline", 
                       "ivh",
                      # "hydrocephalus",
                       "ich.volume.cc",
                       "ivh.volume.cc",
                       "phe.volume.cc",
                       "sp.shift.baseline",
                       "mech.ventilation",
                       "intubation.days", 
                       "extubation.date", # Date of extubation
                       "tracheostomy",
                       "trach.date", # Date of Tracheostomy
                       "evd",
                       "evd.days", # Total EVD days
                       "nsgy.evacuation",
                       "nsgy.evacuation.date",  # Date of Evacuation
                       "mrs.30",
                       "mrs.90",
                      "mrs.180",
                      "mrs.365",
                      "euro.mobility.90",
                      "euro.mobility.180",
                      "euro.mobility.365",
                       "euro.selfcare.90",
                      "euro.selfcare.180",
                      "euro.selfcare.365",
                       "euro.usual.90",
                      "euro.usual.180",
                      "euro.usual.365",
                       "euro.pain.90",
                      "euro.pain.180",
                      "euro.pain.365",
                       "euro.anxiety.90",
                      "euro.anxiety.180",
                      "euro.anxiety.365",
                       "euro.vas.90", # EuroQOL VAS
                       "euro.vas.180",
                       "euro.vas.365",
                       "EQ_INDEX.90", # EQ-5D Utility Index,,
                      "current.location.90",
                      "current.location.180",
                      "current.location.365",
                       "ich.laterality"
                  )

@@

**Importing and wrangling ERICH data**

@@

erich.whole <- erich.whole %>%
  mutate(evd = case_when(
    IT8 == "1" ~ "Yes",
    IT8 == "2" ~ "No",
    str_detect(erich.whole$IT5F_SPEC, "VENTR") ~ "Yes",
  ))

erich.whole <- erich.whole %>%
  mutate(nsgy.evacuation = case_when(
    IT5D == "1" ~ "Yes",
    IT5H == "1" ~ "Yes",
    str_detect(erich.whole$IT5F_SPEC, "CRANI") ~ "Yes",
    str_detect(erich.whole$IT5F_SPEC, "EVAC") ~ "Yes",
    TRUE ~ "No"
      ))

@@

# Adding tracheostomy information

erich.whole <- erich.whole %>%
  mutate(
    tracheostomy = case_when(
      str_detect(erich.whole$IT5F_SPEC, "TRACHEO") ~ "Yes",
      ))

# Variables missing: nihss.baseline, hydrocephalus, phe.volume.cc, sp.shift.baseline, intubation.days

@@

**Importing and wrangling iDEF data**

```{r iDEF}

# Create function to paste together string of formXX.sas7bdat
idef.import <- function(y) {
  dataset <- paste0("idef",".",y)
  filepath <- paste0("./iDEF/form",y,".sas7bdat")
  dataset <- read_sas(filepath)
}

idefdictionary <- read_sas("iDEF/datadictionary.sas7bdat")
idef01 <- idef.import("01")
idef02 <- idef.import("02")
idef03 <- idef.import("03")
idef03c <- idef.import("03c")
idef04 <- idef.import("04")
idef05 <- idef.import("05")
idef06 <- idef.import("06")
idef07 <- idef.import("07")
idef08 <- idef.import("08")
idef10c <- idef.import("10c")
idef11 <- idef.import("11")
idef15 <- idef.import("15")
idef18 <- idef.import("18")
idef19 <- idef.import("19")
idef19c <- idef.import("19c")
idef22 <- idef.import("22")
idef23 <- idef.import("23")
idef24 <- idef.import("24")
idef25 <- idef.import("25")
idef25c <- idef.import("25c")
idef26 <- idef.import("26")
idef27 <- idef.import("27")
idef29 <- idef.import("29")
idef30 <- idef.import("30")
idef32 <- idef.import("32")
idef32c <- idef.import("32c")
idef43 <- idef.import("43")

**Importing and wrangling CLEAR-III data**

```{r CLEAR III}

# Importing Excel files
clear3.demographics <- readxl::read_excel("./CLEAR3/CLEAR III_Publication data_2018.1.30.xlsx", sheet = 1)
clear3.outcomes <- readxl::read_excel("./CLEAR3/CLEAR III_Publication data_2018.1.30.xlsx", sheet = 2)
clear3.180outcomes <- readxl::read_excel("./CLEAR3/CLEAR III_clinicaltrials.gov results_2018.1.30.xlsx", sheet = 1)

# Combining all dataframes into one
clear3.whole <- left_join(clear3.demographics, clear3.outcomes)
clear3.whole <- left_join(clear3.whole, clear3.180outcomes)

# Adding category to indicate study identity
clear3.whole <- clear3.whole %>% mutate(study = "CLEAR III")

clear3.whole <- clear3.whole %>%
  mutate(race = case_when(
    race == "African-American" ~ "Black or African-American",
    race == "Asian" ~ "Asian",
    race == "Hawaiian or Pacific" ~ "Native Hawaiian/Pacific Islander",
    race == "Indian or Alaskan" ~ "American Indian or Alaskan Native",
    race == "Indian or Alaskan;White" ~ "American Indian or Alaskan Native",
    race == "unknown/missing" ~ "Other/Not Reported",
    race == "White" ~ "White"
  ))

clear3.whole <- clear3.whole %>%
  mutate(ethnicity = recode(hispanic_latino,
                            "0" = "Not Hispanic or Latino",
                            "1" = "Hispanic or Latino"
  ))

clear3.whole <- clear3.whole %>%
  mutate(sex = recode(gender,
                         "female" = "Female",
                         "male" = "Male"))

clear3.whole <- clear3.whole %>%
  mutate(ich.laterality = recode(hemisphere,
                                 "left" = "Left",
                                 "right" = "Right",
                                 "unknown/missing" = "NA" ))

clear3.whole <- clear3.whole %>% mutate(ivh = "Yes")

clear3.whole <- clear3.whole %>%
  rename(age = age_at_registration) %>%
  rename(tobacco = tobacco_use) %>%
  rename(sbp.initial = er_present_systolic) %>%
  rename(dbp.initial = er_present_diastolic) %>%
  rename(ich.location.baseline = adjudicated_ich_location) %>%
  rename(gcs.baseline = er_present_gcs_total) %>%
  rename(nihss.baseline = er_present_nihss_total) %>%
  rename(ich.volume.cc = dct_ich_volume_rc) %>%
  rename(ivh.volume.cc = dct_ivh_volume_rc) %>%
  rename(mrs.30 = glasgow_rankin_30) %>%
  rename(mrs.180 = glasgow_rankin_180) %>%
  rename(gose.30 = gose_30) %>%
  rename(gose.180 = gose_180) %>%
  rename(comfort.binary = withdrawal_contribution) %>%
  rename(comfort.date.time = ictus_to_withdrawal_days) %>%
  rename(symptom.onset.to.presentation = ictus_to_presentation)

**Importing and Wrangling MISTIE-II data**

```{r MISTIE II}

# Importing MISTIE II data

mistie2.demographics <- readxl::read_excel("./MISTIE2/MISTIE II_Publication results_2016.4.14_RT_RD updates 2017.10.2.xlsx", sheet = 1)

mistie2.demographics2 <- readxl::read_excel("./MISTIE2/MISTIE II and ICES_Demographics data_2017.10.16.xlsx", sheet = 1)

mistie2.results <- readxl::read_excel("./MISTIE2/MISTIE II and ICES_ctgov results_2017.10.17.xlsx", sheet = 3)

mistie2.demographics3 <- readxl::read_excel("./MISTIE2/ICES_Publication results_2015.12.23_RT_RD updates 2017.10.16.xlsx", sheet = 1)

mistie2.combined.demographics <- full_join(mistie2.demographics, mistie2.demographics3)
mistie2.combined.demographics <- full_join(mistie2.combined.demographics, mistie2.demographics2)

mistie2.wide.results <- mistie2.results %>%
  pivot_wider(names_from = Follow_up_Visit, values_from = c(GOS, eGOS_Score, rankin_score, SIS_Total, SIS_16_Total))

mistie2.whole <- full_join(mistie2.wide.results, mistie2.combined.demographics)

mistie2.whole <- mistie2.whole %>% mutate(study = "MISTIE II")

mistie2.whole <- mistie2.whole %>%
  mutate(ethnicity = case_when(
    Race == "African American not Hispanic" ~ "Not Hispanic or Latino",
    Race == "Hispanic" ~ "Hispanic or Latino",
    Race == "Asian or Pacific Islander" ~ "Not Hispanic or Latino",
    Race == "Caucasian not Hispanic" ~ "Not Hispanic or Latino",
    Race == "Other or unknown" ~ "Not Hispanic or Latino"
  ))

mistie2.whole <- mistie2.whole %>%
  mutate(race = recode(Race,
                       "African American not Hispanic" = "Black or African American",
                       "Asian or Pacific Islander" = "Asian",
                       "Caucasian not Hispanic" = "White",
                       "Hispanic" = "Other/Not Reported",
                       "Other or unknown" = "Other/Not Reported"
  ))

mistie2.whole <- mistie2.whole %>%
  rename(sbp.initial = er_sbp) %>%
  rename(dbp.initial = er_dbp) %>%
  rename(gcs.baseline = enrollment_gcs) %>%
  rename(nihss.baseline = enrollment_nihss_total) %>%
  rename(ich.volume.cc = pre_rand_ichvol) %>%
  rename(ivh.volume.cc = pre_rand_ivhvol) %>%
  rename(gos.365 = GOS_365) %>%
  rename(gos.270 = GOS_270) %>%
  rename(gos.180 = GOS_180) %>%
  rename(gos.90 = GOS_90) %>%
  rename(gos.30 = GOS_30) %>%
  rename(egos.365 = eGOS_Score_365) %>%
  rename(egos.270 = eGOS_Score_270) %>%
  rename(egos.180 = eGOS_Score_180) %>%
  rename(egos.90 = eGOS_Score_90) %>%
  rename(egos.30 = eGOS_Score_30) %>%
  rename(mrs.365 = rankin_score_365) %>%
  rename(mrs.270 = rankin_score_270) %>%
  rename(mrs.180 = rankin_score_180) %>%
  rename(mrs.90 = rankin_score_90) %>%
  rename(mrs.30 = rankin_score_30)

mistie2.whole <- mistie2.whole %>%
  mutate(ivh = recode(any_ivh_prerand,
                      "0" = "No",
                      "1" = "Yes"
  ))

**Importing and Wrangling MISTIE-III data**

```{r MISTIE-III}

# Importing MISTIE III data
mistie3 <- read.table("./MISTIE3/Deidentified_Mistie_III_Wide_Analytic_2019_07_11 (1).csv", sep = ",", header = TRUE)

mistie3 <- mistie3 %>% mutate(study = "MISTIE III")

mistie3 <- mistie3 %>%
  rename(age = age_at_consent) %>%
  rename(race = race_ninds) %>%
  rename(sbp.initial = er_present_systolic) %>%
  rename(dbp.initial = er_present_diastolic) %>%
  rename(gcs.baseline = gcs_total) %>%
  rename(nihss.baseline = nihss_total) %>%
  rename(ich.volume.cc = diagct_ich_volume) %>%
  rename(ivh.volume.cc = diagct_ivh_volume) %>%
  rename(ich.laterality = ich_hemisphere_adj) %>%
  rename(mrs.30 = glasgow_rankin_30) %>%
  rename(mrs.180 = glasgow_rankin_180) %>%
  rename(mrs.365 = glasgow_rankin_365) %>%
  rename(barthel.total.365 = barthel_index_365) %>%
  rename(eq.vas.365 = eq_vas_365) %>%
  rename(time.to.presentation = ictus_to_site_arrival_hours)

mistie3 <- mistie3 %>%
  mutate(race = case_when(
    race == "African-American" ~ "Black of African-American",
    race == "Asian Hawaiian or Pacific" ~ "Asian",
    race == "Indian or Alaskan" ~ "American Indian or Alaskan Native",
    race == "More than one race" ~ "Other/Not Reported",
    race == "White" ~ "White"
  ))

mistie3 <- mistie3 %>%
  mutate(ethnicity = case_when(
         hispanic == "0" ~ "Not Hispanic or Latino",
         hispanic == "1" ~ "Hispanic or Latino"
  ))

mistie3 <- mistie3 %>%
  mutate(sex = case_when(
    male_gender == "0" ~ "Woman",
    male_gender == "1" ~ "Man"
  ))

mistie3 <- mistie3 %>%
  mutate(tobacco = case_when(
    current_smoker == "0" ~ "No",
    current_smoker == "1" ~ "Yes"
  ))

mistie3 <- mistie3 %>%
  mutate(dm2 = case_when(
    diabetes_pmh == "0" ~ "No",
    diabetes_pmh == "1" ~ "Yes"
  ))

mistie3 <- mistie3 %>%
  mutate(htn = case_when(
    htn_pmh == "0" ~ "No",
    htn_pmh == "1" ~ "Yes"
  ))

mistie3 <- mistie3 %>%
  mutate(cad = case_when(
    cvd_pmh == "0" ~ "No",
    cvd_pmh == "1" ~ "Yes"
  ))

mistie3 <- mistie3 %>%
  mutate(comfort = case_when(
    withdrawal_contribution == "0" ~ "No",
    withdrawal_contribution == "1" ~ "Yes"
  ))

mistie3 <- mistie3 %>%
  mutate(euro.mobility.365 = case_when(
    eq5d_mobility_365 == "1" ~ "No Problems",
    eq5d_mobility_365 == "2" ~ "Some Problems",
    eq5d_mobility_365 == "3" ~ "Extreme Problems"
  ))

mistie3 <- mistie3 %>%
  mutate(euro.selfcare.365 = case_when(
    eq5d_selfcare_365 == "1" ~ "No Problems",
    eq5d_selfcare_365 == "2" ~ "Some Problems",
    eq5d_selfcare_365 == "3" ~ "Extreme Problems"
  ))

mistie3 <- mistie3 %>%
  mutate(euro.usual.365 = case_when(
    eq5d_activities_365 == "1" ~ "No Problems",
    eq5d_activities_365 == "2" ~ "Some Problems",
    eq5d_activities_365 == "3" ~ "Extreme Problems"
  ))

mistie3 <- mistie3 %>%
  mutate(euro.pain.365 = case_when(
    eq5d_pain_365 == "1" ~ "No Problems",
    eq5d_pain_365 == "2" ~ "Some Problems",
    eq5d_pain_365 == "3" ~ "Extreme Problems"
  ))

mistie3 <- mistie3 %>%
  mutate(euro.anxiety.365 = case_when(
    eq5d_anxiety_365 == "1" ~ "No Problems",
    eq5d_anxiety_365 == "2" ~ "Some Problems",
    eq5d_anxiety_365 == "3" ~ "Extreme Problems"
  ))

mistie3 <- mistie3 %>%
  mutate(current.location.30 = case_when(
    location_d30 == "Acute - Hospital" ~ "Hospital",
    location_d30 == "Acute - ICU" ~ "Hospital",
    location_d30 == "Dead" ~ "Dead",
    location_d30 == "Home" ~ "Home",
    location_d30 == "Rehab" ~ "Rehabilitation",
    location_d30 == "LTCF" ~ "Skilled Nursing Facility",
    location_d30 == "Hospice" ~ "Hospice",
  ))

mistie3 <- mistie3 %>%
  mutate(current.location.180 = case_when(
    location_d180 == "Acute - Hospital" ~ "Hospital",
    location_d180 == "Acute - ICU" ~ "Hospital",
    location_d180 == "Dead" ~ "Dead",
    location_d180 == "Home" ~ "Home",
    location_d180 == "Rehab" ~ "Rehabilitation",
    location_d180 == "LTCF" ~ "Skilled Nursing Facility",
    location_d180 == "Hospice" ~ "Hospice",
  ))

mistie3 <- mistie3 %>%
  mutate(current.location.365 = case_when(
    location_d365 == "Acute - Hospital" ~ "Hospital",
    location_d365 == "Acute - ICU" ~ "Hospital",
    location_d365 == "Dead" ~ "Dead",
    location_d365 == "Home" ~ "Home",
    location_d365 == "Rehab" ~ "Rehabilitation",
    location_d365 == "LTCF" ~ "Skilled Nursing Facility",
    location_d365 == "Hospice" ~ "Hospice",
  ))

**Combining Datasets**
